---
title: "p8105_hw2_yy3662"
author: "Yunyi Yang"
date: "2025-09-29"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1

```{r}
library(tidyverse)
library(skimr)
p1task1 <- read.csv(file = "./Data/fivethirtyeight_datasets/pols-month.csv",
                  ,na = c(".", "NA", "")) |>
  separate(mon, c("year", "month", "day"), convert = TRUE) |>
  mutate(month = month.name[month]) |>
  pivot_longer(cols = c(prez_dem, prez_gop), names_to = "parties", values_to = "president") |>
  select(-day)



p1task2 <- read.csv(file = "./Data/fivethirtyeight_datasets/snp.csv",
                  ,na = c(".", "NA", ""))
p1task2$date <- as.Date(p1task2$date, "%d/%m/%y")
p1task2$date <-as.character(p1task2$date)
p1task2 <- separate(p1task2, date, c("year", "month", "day"), convert = TRUE) |>
  mutate(p1task2, month = month.name[month]) |>
  select(-date) # I find date is unnecessary in final dataset
p1task2$year[which(p1task2$year > 2025)] <- p1task2$year[which(p1task2$year > 2025)] - 100L
# because some dates are incorrect because of basic logic of as.Date, so we convert some years like 2065 back to 1965.


p1task3 <- read.csv(file = "./Data/fivethirtyeight_datasets/unemployment.csv",
                  ,na = c(".", "NA", "")) |>
  pivot_longer(Jan:Dec, names_to = "month", values_to = "unemployment") |>
  mutate(month = month.name[match(tolower(month), tolower(month.abb))]) |>
  rename(year = "Year")

join1 <- full_join(p1task1, p1task2, by = c("year", "month"))

join2 <- left_join(join1, p1task3, by = c("year", "month"))
skimr::skim(join2)
summary(join2)
```

pols-month which names p1task1 contains 822 observations of 9 variables related to the number of national politicians who are democratic or republican at any given time:
mon: date of the count
prez_gop: indicator of whether the president was republican on the associated date (1 = yes, 0 = no)
gov_gop: the number of republican governors on the associated date
sen_gop: the number of republican senators on the associated date
rep_gop: the number of republican representatives on the associated date
prez_dem: indicator of whether the president was democratic on the associated date (1 = yes, 0 = no)
gov_dem: the number of democratic governors on the associated date
sen_dem: the number of democratic senators on the associated date
rep_dem: the number of democratic representatives on the associated date

snp which names p1task2 contains 787 observations of 2 variables related to Standard & Poor’s stock market index (S&P), often used as a representative measure of stock market as a whole:
date: the date of the observation
close: the closing values of the S&P stock index on the associated date

unemployment which names p1task3 contains 3 variables:
year: year of measurement
month: month of measurement
unemployment: the percentage of unemployment of specific year and month

In our final dataset, we have 2782 observations and 13 variables, where contains year, month, day, gov_gop, sen_gop, rep_gop, gov_dem, sen_dem, rep_dem, parties which shows two options of parties, and president shows which party is president from at that year, close, unemployment. The year range is from 1947-01 to 2015-06.

# Problem 2
```{r}
library(readxl)
mrtrash <- read_excel("./Data/202409 Trash Wheel Collection Data.xlsx", sheet = 1, range = "A2:N653")
mrtrash <- janitor::clean_names(mrtrash) |>
  mutate(mrtrash$name, name = "mrtrash")
mrtrash$sports_balls <- as.integer(round(mrtrash$sports_balls))

professor <- read_excel("./Data/202409 Trash Wheel Collection Data.xlsx", sheet = 2, range = "A2:M120")
professor <- janitor::clean_names(professor) |>
  mutate(professor$name, name = "professor")

Gwynnda <- read_excel("./Data/202409 Trash Wheel Collection Data.xlsx", sheet = 4, range = "A2:L265")
Gwynnda <- janitor::clean_names(Gwynnda) |>
  mutate(Gwynnda$name, name = "Gwynnda")

mrtrash$year <- as.numeric(mrtrash$year) #mrtrash's year variable is not numeric as professor and Gwynnda
join3 <- bind_rows(mrtrash, professor, Gwynnda)
skimr::skim(join3)
summary(join3)

sum(filter(join3, name == "mrtrash")$weight_tons)
sum(filter(join3, name == "Gwynnda", year == "2022", month == "June")$cigarette_butts)
```
Our final dataset has 1032 observations and 15 variables, which includes "dumpster" series number, "month", "year" and "date" that the dumpster was collected, and "weight_tons" which means total weights of trash in the dumpster, "volumn_cubic_yards" which means volumn in cubic yards, and amount of different trash includes: "plastic_bottles", "polystyrene", cigarette_butts", "glass_bottles", "plastic_bags", "wrappers", "sports_balls", and "homes_powered" describes how many homes are powered by these energy from trash, and "name" indicates which trash wheel collected these trash. 
The total weight of Mr. Trash Wheel is 2091.18 tons, and total number of cigarettes butts in 2022.6 from Gwynnda is 18120.

# Problem 3
```{r}
```